Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-02-06 content.tex
EOD

We show the usage of \verb;pirouette; by gradually introducing
its features.
First, to get an idea of the baseline error, 
we measure the Bayesian inference error when we start from a phylogeny 
generated under a known and standard tree model.
Second, to establish that the true tree prior is indeed the best, 
we also measure the inference error made by a
best candidate inference model.
Lastly, we quantify the impact of the tree prior in the Bayesian inference. 
We do so by separating the baseline error from the complete error 
running the pipeline starting from a tree generated by a non-standard tree model.

@@

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single, 
    label = {lst:create_yule_tree}, 
    caption = {
      Create a Yule tree. 
      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
    }
  ]
phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
\end{lstlisting}

\begin{figure}[ht]
  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/true_tree.png}
  \caption{The Yule tree, as created by Listing~\ref{lst:create_yule_tree}.}
  \label{fig:yule_tree}
\end{figure}

@@

We will use the same tree as generated in \ref{lst:create_yule_tree}, as well 
as the same alignment parameters as shown 
in Listing~\ref{lst:create_alignment_params}.

@@

We start from the tree generated by a diversity-dependent model,
having six taxa and a crown age of ten:
\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
# NEW CODE HERE, 
# depends on https://github.com/richelbilderbeek/pirouette_article/issues/75
#
# phylogeny  <- ape::read.tree(
#   text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
# )
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/true_tree.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
    The scale overlaps with the lower branch, which is the default 
    behavior for the plotting function used.
  }
\end{figure}

>>>>

We show the usage of \verb;pirouette; on an tree generated by a non-standard 
tree model.

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single, 
    label = {lst:create_dd_tree}, 
    caption = {
      Create a Yule tree. 
      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
    }
  ]
phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
\end{lstlisting}

by gradually introducing
its features.
First, to get an idea of the baseline error, 
we measure the Bayesian inference error when we start from a phylogeny 
generated under a known and standard tree model.
Second, to establish that the true tree prior is indeed the best, 
we also measure the inference error made by a
best candidate inference model.
Lastly, we quantify the impact of the tree prior in the Bayesian inference. 
We do so by separating the baseline error from the complete error 
running the pipeline starting from a tree generated by a non-standard tree model.

@@

\begin{figure}[ht]
  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/true_tree.png}
  \caption{The Yule tree, as created by Listing~\ref{lst:create_dd_tree}.}
  \label{fig:yule_tree}
\end{figure}

@@

We will use the same tree as generated in \ref{lst:create_dd_tree}, as well 
as the same alignment parameters as shown 
in Listing~\ref{lst:create_alignment_params}.

@@

We start from the tree generated by a diversity-dependent model,
having six taxa and a crown age of ten:
\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/true_tree.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
    The scale overlaps with the lower branch, which is the default 
    behavior for the plotting function used.
  }
\end{figure}
